Stochastic Wind Power Forecasting

### 2020-04-10 LaTeX_Code/Paper/paper_1line.tex
moments beta

Now, looking at the normalized wind power generation forecast process $X$, modeled as solution to the It\^{o} stochastic differential equation (\ref{eq:main}) with the drift specified in (\ref{drift:meanrev-derivtrack}), it is straightforward to check that $\mathbb{E} \left[X_t\right] = p_t$, given $  \mathbb{E} X_0 = p_0$. The application of  It\^{o}'s lemma on $g(X_t, t) = X_t e^{\int_{0}^{t} \theta_s ds}$, leads to 
\begin{equation*}
d \left( X_t e^{\int_{0}^{t} \theta_s ds} \right) = e^{\int_{0}^{t} \theta_s ds} (\dot{p}_t + \theta_t p_t ) dt  + e^{\int_{0}^{t} \theta_s ds}\, b (X_t; p_t, \dot{p}_t, \bm{\theta} ) \,dW_t\,,
\end{equation*}
whose integral form is
\begin{equation}
e^{\int_{0}^{t} \theta_s ds}  X_t - X_0 = \int_{0}^{t}  (\dot{p}_s + \theta_s p_s ) e^{\int_{0}^{s} \theta_u du} ds + \int_{0}^{t}  b (X_s; p_s, \dot{p}_s, \bm{\theta} )  e^{\int_{0}^{s} \theta_u du}\, d W_s\,.   \label{eq:tomeanX}
\end{equation} 
Taking expectation on (\ref{eq:tomeanX}) we obtain
\begin{align}
\mathbb{E} X_t & = e^{ - \int_{0}^{t} \theta_s ds} \left(  \mathbb{E} X_0 +   \int_{0}^{t}  (\dot{p}_s + \theta_s p_s ) e^{\int_{0}^{s} \theta_u du} ds   \right) \nonumber \\ 
& =  e^{ - \int_{0}^{t} \theta_s ds} \left(  \mathbb{E} X_0 +   \int_{0}^{t}  \dot{p}_s e^{\int_{0}^{s} \theta_u du} ds +  \big[ p_s \, e^{\int_{0}^{s} \theta_u du} \big]^{t}_0  - \int_{0}^{t}  \dot{p}_s e^{\int_{0}^{s} \theta_u du} ds  \right) \nonumber \\
& =  e^{ - \int_{0}^{t} \theta_s ds} \left(  \mathbb{E} X_0 + p_t \, e^{ \int_{0}^{t} \theta_s ds} - p_0 \right) = p_t \,.
\end{align}

@@

Closed-form solutions to initial-boundary value problem for time-in\-ho\-mo\-geneous diffusions can be obtained only in a few cases. Besides, in our case solving numerically (\ref{eq:fpk}) for the transition densities of the process $V$ at every transition step is computationally expensive. 
Therefore, under the likelihood-based inferential paradigma, many techniques have been devised to obtain approximate maximum likelihood estimates for the unknown parameters of continuous-time SDE models with discrete observations. Parametric estimation problems for diffusion processes sampled at discrete times are presented in \autocite[Chapter 3]{iacus1}, and a survey of estimation methods for the parameter vector of the general one-dimensional, time-homogeneous SDE from a single sample of observations at discrete times is presented in \autocite{hurn}.

@@

Using It\^o's lemma on , we arrive at the following iterative ODEs for the state dependent diffusion formulation (\ref{VtSDE})
\begin{equation}
\frac{d \mathbb{E}[ V^k_t]}{dt} = - k \theta_t \mathbb{E}[ V^k_t] + \frac{k(k-1)}{2} \mathbb{E}[ V^{k-2}_t  b(V^k_t;\theta_t, \alpha)]
\end{equation}

For $t\in [t_{n-1}, t]$, the first two moments are given by
\begin{align}
\frac{d m_1 (t)}{dt} &= - m_1(t)\theta_t   \nonumber \\
\frac{d m_2 (t)}{dt} &=  -2 m_2(t)\theta_t(1+\alpha) + 2\alpha\theta_t m_1(t)(1-2p_t)  \nonumber \\
&\quad + 2 \alpha\theta_t p_t (1-p_t) \,,
\end{align}
with initial conditions $m_1(t_{n-1})= v_{n-1}$ and $m_2(t_{n-1})= v_{n-1}^2 \,.$

@@

A suitable candidate for a proxy transition density is a Beta probability distribution as it is compactly supported and can morph into symmetric and asymmetric shapes. {\color{red} Move this sentence to next section.}

\subsubsection*{ Moment Matching}
To approximate the transition densities of  the process $V_t$ by a Beta distribution, we match its moments with the shape parameters $\xi_1, \xi_2$ of a Beta proxy density on $[-1,1]$. The any moment of the process $V_t$ is  given by solving the corresponding ODE system  for the $k^{th}$  moment,

The shape parameters are given by
\begin{equation}
\xi_1 = - \frac{(1+\mu_t )(\mu_t^2 + \sigma_t^2 -1)}{2 \sigma_t^2}\,, \quad \xi_2=  \frac{(\mu_t-1 )(\mu_t^2 + \sigma_t^2 -1)}{2 \sigma_t^2} \,, \label{param_transformed_beta}
\end{equation}
where $\mu_t = m_1 (t)$ and $\sigma_t^2= m_2 (t)- m_1 (t)^2\,.$

>>>>

Now, looking at the normalized wind power generation forecast process $X$, modeled as solution to the It\^{o} stochastic differential equation (\ref{eq:main}) with the drift specified in (\ref{drift:meanrev-derivtrack}), it is straightforward to check that $\mathbb{E} \left[X_t\right] = p_t$, given $  \mathbb{E} X_0 = p_0$. The application of  It\^{o}'s lemma on $g(X_t, t) = X_t e^{\int_{0}^{t} \theta_s ds}$, leads to 
\begin{equation*}
d \left( X_t e^{\int_{0}^{t} \theta_s ds} \right) = e^{\int_{0}^{t} \theta_s ds} (\dot{p}_t + \theta_t p_t ) dt  + e^{\int_{0}^{t} \theta_s ds}\, b (X_t; p_t, \dot{p}_t, \bm{\theta} ) \,dW_t\,,
\end{equation*}
whose integral form is
\begin{equation}
e^{\int_{0}^{t} \theta_s ds}  X_t - X_0 = \int_{0}^{t}  (\dot{p}_s + \theta_s p_s ) e^{\int_{0}^{s} \theta_u du} ds + \int_{0}^{t}  b (X_s; p_s, \dot{p}_s, \bm{\theta} )  e^{\int_{0}^{s} \theta_u du}\, d W_s\,.   \label{eq:tomeanX}
\end{equation} 
Taking expectation on (\ref{eq:tomeanX}) we obtain
\begin{align}
\mathbb{E} X_t & = e^{ - \int_{0}^{t} \theta_s ds} \left(  \mathbb{E} X_0 +   \int_{0}^{t}  (\dot{p}_s + \theta_s p_s ) e^{\int_{0}^{s} \theta_u du} ds   \right) \nonumber \\ 
& =  e^{ - \int_{0}^{t} \theta_s ds} \left(  \mathbb{E} X_0 +   \int_{0}^{t}  \dot{p}_s e^{\int_{0}^{s} \theta_u du} ds +  \big[ p_s \, e^{\int_{0}^{s} \theta_u du} \big]^{t}_0  - \int_{0}^{t}  \dot{p}_s e^{\int_{0}^{s} \theta_u du} ds  \right) \nonumber \\
& =  e^{ - \int_{0}^{t} \theta_s ds} \left(  \mathbb{E} X_0 + p_t \, e^{ \int_{0}^{t} \theta_s ds} - p_0 \right) = p_t \,. \label{eq:meanX}
\end{align}

@@

Closed-form solutions to initial-boundary value problem for time-in\-ho\-mo\-geneous diffusions can be obtained only in a few cases. Besides, in our case solving numerically (\ref{eq:fpk}) for the transition densities of the process $V$ at every transition step is computationally expensive. 
Therefore, under the likelihood-based inferential paradigma, many techniques have been devised to obtain approximate maximum likelihood estimates for the unknown parameters of continuous-time SDE models with discrete observations. 
 {\color{red} Rewrite: Parametric estimation problems for diffusion processes sampled at discrete times are presented in \autocite[Chapter 3]{iacus1}, and a survey of estimation methods for the parameter vector of the general one-dimensional, time-homogeneous SDE from a single sample of observations at discrete times is presented in \autocite{hurn}.}

@@

From (\ref{eq:meanX}), we have $m_1(t) \equiv \mathbb{E} V_t = e^{- \int_{t_{j,i-1}}^t \theta_s ds} \,\mathbb{E} V_{t_{j,i-1}}$, for any $t\in [t_{j,i-1}, t_{j, i})$, $j = 1, \ldots, M$ and $i = 1, \ldots, N$ .

For $m \geq 2$, using It\^o's lemma on $g(V_t) = V_t^m$, we obtain
% arrive at the following iterative ODEs for the state dependent diffusion formulation (\ref{VtSDE})
\begin{align}
d(V_t^m) & = \left( - m \theta_t V_t^m + \frac{1}{2} m (m -1) V_t^{m -2} 2 \alpha \theta_0 (V_t + p_t) (1 - V_t - p_t) \right) dt  \nonumber \\
& + m V_t^{m-1} 2 \alpha \theta_0 (V_t + p_t) (1 - V_t - p_t) dW_t \,, \nonumber
\end{align}
from which we derive
\begin{equation}
\frac{d \mathbb{E}[ V^m_t]}{dt} = - m \theta_t \mathbb{E}[ V^m_t] + m (m-1) \alpha \theta_0  \mathbb{E}[ - V_t^m + (1 - 2 p_t) V_t^{m-1} + p_t (1 -p_t) V_t^{m-2} ] \,.
\end{equation}

For any $t\in [t_{j,i-1}, t_{j, i})$, the first two moments of $V$, $m_1(t)$ and $m_2(t) \equiv \mathbb{E}[V_t^2]$, can be computed by solving the following system
\begin{equation}
  \left\{
  \begin{array}{@{}rl@{}}
    \frac{d m_1 (t)}{dt} \!\!\!&=  - m_1(t)\theta_t   \\
   \frac{d m_2 (t)}{dt}  \!\!\!&=  -2 (\theta_t +\alpha \theta_0) m_2(t) + 2 \alpha\theta_0 (1-2p_t)  m_1(t) + 2 \alpha\theta_0 p_t (1-p_t) 
 \end{array}\right.  \label{Vtmom}
\end{equation}
with initial conditions $m_1(t_{j,i-1})= v_{j, i-1}$ and $m_2(t_{j,i-1})= v_{j, i-1}^2 \,.$

\subsubsection*{ Moment Matching}
A suitable candidate for a surrogate transition density of $V$ is a Beta distribution on a compact interval parameterized by two positive shape real parameters, $\xi_1, \xi_2$.

To approximate the transition densities of  the process $V$ using a Beta distribution, we equal the first two central moments of $V$ with the corresponding moments of the Beta surrogate distribution on $[-1 + \epsilon,1 - \epsilon]$ with shape parameters $\xi_1, \xi_2$ .

The shape parameters are given by
\begin{equation}
\xi_1 = - \frac{(\mu_t + 1 - \epsilon)(\mu_t^2 + \sigma_t^2 - (1- \epsilon)^2)}{2 (1 - \epsilon) \sigma_t^2}\,, \quad \xi_2=  \frac{(\mu_t-1 + \epsilon )(\mu_t^2 + \sigma_t^2 - (1- \epsilon)^2)}{2 (1 - \epsilon) \sigma_t^2} \,, \label{param_transformed_beta}
\end{equation}
where $\mu_t = m_1 (t)$ and $\sigma_t^2= m_2 (t)- m_1 (t)^2\,.$ \\

@@
